---
title: "environment_fg"
author: "Katja Kozjek"
output: 
  rmarkdown::html_document:
    code_folding: 'hide'
    toc: true
    toc_float: true
    smart: true
    number_sections: false
    highlight: tango
    self_contained: true
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r wrap-hook, include=FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

#load packages 

```{r setting the environment, message=F, warning=F}

library(car)
library(lmerTest)
library(lmtest)
library(multcomp)
library(ggplot2)
library(wesanderson)
library(emmeans)

```

***IMPORTANT: drought treatment, and country are taken as factors***
***SOC as continuous variable*** 

***because 1. field pairs do not always differ in SOC, 2. “low” and “high” present very different values not just among countries, but also within countries and 3. using SOC as a continuous gives us an idea about the actual values that matter, field pairs are accounted in the random structure***

# Water content 

## volumetric water content

```{r volumetric water content, message=FALSE, warning=FALSE}

water_content <- read.csv("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/metafile/water_content.csv")

str(water_content)

data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

#volumetric soil water content (bulk density*water content)

water_summary <- data_summary(water_content, varname="VWC", 
                 groupnames=c("treatment", "country"))

water_summary$treatment<-factor(water_summary$treatment, levels=c("R", "RC", "C"))
water_summary$country<-factor(water_summary$country, levels=c("SE", "DE", "ES"),labels = c("Sweden", "Germany", "Spain"))

#plot volumetric soil water content 
p_water <- ggplot(water_summary, aes(treatment, VWC)) +
    geom_errorbar(
    aes(ymin = VWC-sd, ymax = VWC+sd, color = treatment),
    position = position_dodge(0.3), width = 0.2
    ) +
    geom_point(aes(color = treatment), position = position_dodge(0.3)) + facet_grid(.~country) + theme_bw() + xlab("Drought treatment") + ylab("Soil water content [%Vol]") + scale_color_manual(values=wes_palette(n=3, name="Moonrise2")) + labs(color = "Drought treatment")

print(p_water)

#ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/vWC.jpeg", width = 20, height = 10, units = "cm")

```

## gravimetric water content

```{r gravimetric water content, message=FALSE, warning=FALSE}

#gravimetric soil water content
gwater_summary <- data_summary(water_content, varname="WC", 
                 groupnames=c("treatment", "country"))

gwater_summary$treatment<-factor(gwater_summary$treatment, levels=c("R", "RC", "C"))
gwater_summary$country<-factor(gwater_summary$country, levels=c("SE", "DE", "ES"),labels = c("Sweden", "Germany", "Spain"))

#plot gravimetric soil water content 
p_gwater <- ggplot(gwater_summary, aes(treatment, WC)) +
    geom_errorbar(
    aes(ymin = WC-sd, ymax = WC+sd, color = treatment),
    position = position_dodge(0.3), width = 0.2
    ) +
    geom_point(aes(color = treatment), position = position_dodge(0.3)) + facet_grid(.~country) + theme_bw() + xlab("Drought treatment") + ylab("Soil water content [%]") + scale_color_manual(values=wes_palette(n=3, name="Moonrise2")) + labs(color = "Drought treatment")

print(p_gwater)

#ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/gWC.jpeg", width = 20, height = 10, units = "cm")

```

## WC/WHC ratio

```{r soil moisture reduction, message=FALSE, warning=FALSE}

#soil moisture reduction (WC/WHC)
moisture_summary <- data_summary(water_content, varname="reduction", 
                 groupnames=c("treatment", "country"))

moisture_summary$treatment<-factor(moisture_summary$treatment, levels=c("R", "RC", "C"))
moisture_summary$country<-factor(moisture_summary$country, levels=c("SE", "DE", "ES"),labels = c("Sweden", "Germany", "Spain"))

#plot soil moisture reduction
p_moisture <- ggplot(moisture_summary, aes(treatment, reduction)) +
    geom_errorbar(
    aes(ymin = reduction-sd, ymax = reduction+sd, color = treatment),
    position = position_dodge(0.3), width = 0.2
    ) +
    geom_point(aes(color = treatment), position = position_dodge(0.3)) + facet_grid(.~country) + theme_bw() + xlab("Drought treatment") + ylab("WC/WHC") + scale_color_manual(values=wes_palette(n=3, name="Moonrise2")) + labs(color = "Drought treatment")

print(p_moisture)

#ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/WC_WHC_ratio.jpeg", width = 20, height = 10, units = "cm")

```

# Soil carbon content 

## organic carbon 

```{r organic carbon, message=FALSE, warning=FALSE}

metafile <- read.csv("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/metafile_relaxed.csv", sep=",", row.names=1)

#organic carbon content as continuous 
carbon_summary <- data_summary(metafile, varname="organic.C", 
                 groupnames=c("treatment", "country"))

carbon_summary$treatment<-factor(carbon_summary$treatment, levels=c("R", "RC", "C"))
carbon_summary$country<-factor(carbon_summary$country, levels=c("SE", "DE", "ES"),labels = c("Sweden", "Germany", "Spain"))

#plot % of SOC in high and low
p_organicC <- ggplot(carbon_summary, aes(treatment, organic.C)) +
    geom_errorbar(
    aes(ymin = organic.C-sd, ymax = organic.C+sd, color = treatment),
    position = position_dodge(0.3), width = 0.2
    ) +
    geom_point(aes(color = treatment), position = position_dodge(0.3)) + facet_grid(.~country) + theme_bw() + xlab("Drought treatment") + ylab("Soil organic carbon [%]") + scale_color_manual(values=wes_palette(n=3, name="Moonrise2")) + labs(color = "Drought treatment")

print(p_organicC) 

#ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/organicC.jpeg", width = 20, height = 10, units = "cm")

```

## inorganic C

```{r inorganic carbon, message=FALSE, warning=FALSE}

#inorganic carbon content
inorg_carbon_summary <- data_summary(metafile, varname="inorganic.C", 
                 groupnames=c("treatment", "country"))

inorg_carbon_summary$treatment<-factor(inorg_carbon_summary$treatment, levels=c("R", "RC", "C"))
inorg_carbon_summary$country<-factor(inorg_carbon_summary$country, levels=c("SE", "DE", "ES"),labels = c("Sweden", "Germany", "Spain"))

#plot % inorganic C 
p_inorganicC <- ggplot(inorg_carbon_summary, aes(treatment, inorganic.C)) +
    geom_errorbar(
    aes(ymin = inorganic.C-sd, ymax = inorganic.C+sd, color = treatment),
    position = position_dodge(0.3), width = 0.2
    ) +
    geom_point(aes(color = treatment), position = position_dodge(0.3)) + facet_grid(.~country) + theme_bw() + xlab("Drought treatment") + ylab("Soil inorganic carbon [%]") + scale_color_manual(values=wes_palette(n=3, name="Moonrise2")) + labs(color = "Drought treatment")

print(p_inorganicC)

#ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/inorganicC.jpeg", width = 20, height = 10, units = "cm")

```

# pH value

```{r pH, message=FALSE, warning=FALSE}

#pH value
ph_summary <- data_summary(metafile, varname="pH", 
                 groupnames=c("treatment", "country"))

ph_summary$treatment<-factor(ph_summary$treatment, levels=c("R", "RC", "C"))
ph_summary$country<-factor(ph_summary$country, levels=c("SE", "DE", "ES"),labels = c("Sweden", "Germany", "Spain"))

#plot pH value
p_pH <- ggplot(ph_summary, aes(treatment, pH)) +
    geom_errorbar(
    aes(ymin = pH-sd, ymax = pH+sd, color = treatment),
    position = position_dodge(0.3), width = 0.2
    ) +
    geom_point(aes(color = treatment), position = position_dodge(0.3)) + facet_grid(.~country) + theme_bw() + xlab("Drought treatment") + ylab("pH value") + scale_color_manual(values=wes_palette(n=3, name="Moonrise2")) + labs(color = "Drought treatment") 

print(p_pH)

#ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/pH.jpeg", width = 20, height = 10, units = "cm")

```

# Mixed models 

## model 1 
+ https://ademos.people.uic.edu/Chapter18.html

```{r model 1, message=FALSE, warning=FALSE}

#model for VWC
#carbon as actual value (continuous), as covariate
#plot gets an ID, from 1-30, explicit nesting -> take it as random effect
#ignore field pair

mod1 <- lmer(VWC~country*treatment + SOC + (1|plot) , data=water_content, REML = TRUE)
#test significance of the model
anova_mod1 <- anova(mod1, type = "III")
anova_mod1 #treatment, country effect, slight carbon effect
summary(mod1)

#Check assumptions of the model
qqnorm(resid(mod1)); qqline(resid(mod1)) #Normal distr residuals
#plot(mod1) #Homogeneity (fitted vs residuals)

#Check residuals for each variable (independence and homogeneity)
#plot(resid(mod1)~water_content$country) 
#plot(resid(mod1)~water_content$treatment)

#check the contrasts
summary(glht(mod1, linfct = mcp(treatment = "Tukey")), test = adjusted("holm")) #contrasts between R and C, RC and R
summary(glht(mod1, linfct = mcp(country = "Tukey")), test = adjusted("holm")) #slight contrast between SE and ES

#post hoc test
#Multiple comparison 
m1.emm_country <- emmeans(mod1, ~ country)
m1.emm.res_country <- pairs(m1.emm_country)
m1.emm.res_country #Results are averaged over the levels of: treatment
#DE-SE, ES-SE

plot(m1.emm.res_country, comparisons = TRUE)

m1.emm_treat <- emmeans(mod1, ~ treatment)
m1.emm.res_treat <- pairs(m1.emm_treat)
m1.emm.res_treat #Results are averaged over the levels of: country
#C-R

plot(m1.emm.res_treat, comparisons = TRUE)

```
***This is the final model***

## model 2

```{r model 2, message=FALSE, warning=FALSE}

#global model for VWC
#carbon as actual value (continuous)
#fixed effects 
#plot gets an ID, from 1-30 -> take it as random effect
#ignore field pair

mod2<- lmer(VWC~treatment+country+SOC + (1|plot), data=water_content, REML = TRUE)
#test significance of the model
anova_mod2 <- anova(mod2, type="III")
anova_mod2
summary(mod2)

#Check assumptions of the model
qqnorm(resid(mod2)); qqline(resid(mod2)) #Normal distr residuals
#plot(mod2) #Homogeneity (fitted vs residuals)
#Check residuals for each variable (independence and homogeneity)
#plot(resid(mod2)~water_content$country) 
#plot(resid(mod2)~water_content$treatment)

#check the contrasts
summary(glht(mod2, linfct = mcp(treatment = "Tukey")), test = adjusted("holm"))
summary(glht(mod2, linfct = mcp(country = "Tukey")), test = adjusted("holm"))

```

## model 3 

```{r model 3, message=FALSE, warning=FALSE}

#model for WC/WHC ratio
#exclude carbon
#exclude interactions 
#plot as random effect

lmer_WC1 <- lmer(reduction~country+treatment + (1|plot), data=water_content, REML = TRUE)
#test significance of the model
anova_WC1 <- anova(lmer_WC1, type="III")
anova_WC1
summary(lmer_WC1)

#Check assumptions of the model
qqnorm(resid(lmer_WC1)); qqline(resid(lmer_WC1)) #Normal distr residuals
#plot(lmer_WC1) #Homogeneity (fitted vs residuals)
#Check residuals for each variable (independence and homogeneity)
#plot(resid(lmer_WC1)~water_content$country) 
#plot(resid(lmer_WC1)~water_content$treatment)

#check the contrasts
summary(glht(lmer_WC1, linfct = mcp(treatment = "Tukey")), test = adjusted("holm"))
summary(glht(lmer_WC1, linfct = mcp(country = "Tukey")), test = adjusted("holm"))

```

## model 4

```{r model 4, message=FALSE, warning=FALSE}

#model for WC/WHC ratio
#exclude carbon
#include interactions 
#plot as random effect

lmer_WC2 <- lmer(reduction~treatment*country + (1|plot), data=water_content, REML = TRUE)
#test significance of the model
anova_WC2 <- anova(lmer_WC2, type="III")
anova_WC2
summary(lmer_WC2)

#Check assumptions of the model
qqnorm(resid(lmer_WC2)); qqline(resid(lmer_WC2)) #Normal distr residuals
#plot(lmer_WC2) #Homogeneity (fitted vs residuals)
#Check residuals for each variable (independence and homogeneity)
#plot(resid(lmer_WC2)~water_content$country) 
#plot(resid(lmer_WC2)~water_content$treatment)

#check the contrasts
summary(glht(lmer_WC2, linfct = mcp(treatment = "Tukey")), test = adjusted("holm"))
summary(glht(lmer_WC2, linfct = mcp(country = "Tukey")), test = adjusted("holm"))

```

## model VWC each country

```{r model each country, message=FALSE, warning=FALSE}

#Sweden
sampledata_SE <- read.csv("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/metafile/sampledata_SE.csv", row.names = 1)

lmer_WC_SE <- lmer(VWC~treatment+organic.C + (1|field_pair), data=sampledata_SE, REML = TRUE)
#test significance of the model
anova_WC_SE <- anova(lmer_WC_SE, type="III")
anova_WC_SE
summary(lmer_WC_SE) 

#check the contrasts
summary(glht(lmer_WC_SE, linfct = mcp(treatment = "Tukey")), test = adjusted("holm"))

#Germany
sampledata_DE <- read.csv("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/metafile/sampledata_DE.csv", row.names = 1)

lmer_WC_DE <- lmer(VWC~treatment+organic.C + (1|field_pair), data=sampledata_DE, REML = TRUE)
#test significance of the model
anova_WC_DE <- anova(lmer_WC_DE, type="III")
anova_WC_DE
summary(lmer_WC_DE)

#check the contrasts
summary(glht(lmer_WC_DE, linfct = mcp(treatment = "Tukey")), test = adjusted("holm"))

#Spain
sampledata_ES <- read.csv("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/metafile/sampledata_ES.csv", row.names = 1)

lmer_WC_ES <- lmer(VWC~treatment+organic.C + (1|field_pair), data=sampledata_ES, REML = TRUE)
#test significance of the model
anova_WC_ES <- anova(lmer_WC_ES, type="III")
anova_WC_ES
summary(lmer_WC_ES)

#check the contrasts
summary(glht(lmer_WC_ES, linfct = mcp(treatment = "Tukey")), test = adjusted("holm"))

```

# Regression
http://www.sthda.com/english/wiki/ggplot2-scatter-plots-quick-start-guide-r-software-and-data-visualization

## regression VWC-SOC country

```{r simple regression water-SOC country, message=FALSE, warning=FALSE}

p_water_carbon_country <- ggplot(water_content, aes(x=SOC, y=VWC, color=country), size=6) + geom_point() + geom_smooth(method=lm) + scale_color_manual(values = c("#9C964A", "#B6854D", "#46ACC8")) + labs(colour = "Country") + xlab("Soil organic carbon [%]") + ylab("Soil water content [%Vol]")

print(p_water_carbon_country)

#ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/regression_water_carbon.jpeg", width = 20, height = 10, units = "cm")

```

## regression VWC-SOC treatment

```{r regression water-SOC treatment, message=FALSE, warning=FALSE}

p_water_carbon_treat <- ggplot(water_content, aes(x=SOC, y=VWC, color=treatment), size=6) + geom_point() + geom_smooth(method=lm) + scale_color_manual(values=wes_palette(n=3, name="Moonrise2")) + labs(colour = "Treatment") + xlab("Soil organic carbon [%]") + ylab("Soil water content [%Vol]")

print(p_water_carbon_treat)

#ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/regression_water_carbon_treat.jpeg", width = 20, height = 10, units = "cm")

```

## regression WHC-SOC country

```{r regression WHC-SOC country, message=FALSE, warning=FALSE}

p_whc_carbon <- ggplot(water_content, aes(x=SOC, y=WHC, color=country), size=6) + geom_point() + geom_smooth(method=lm) + scale_color_manual(values = c("#9C964A", "#B6854D", "#46ACC8")) + labs(colour = "Country") + xlab("Soil organic carbon [%]") + ylab("Water holding capacity [%]")

print(p_whc_carbon)

#ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/regression_whc_carbon.jpeg", width = 20, height = 10, units = "cm")

```

## regression WHC-SOC treatment

```{r regression WHC-SOC treatment, message=FALSE, warning=FALSE}

p_whc_carbon_treat <- ggplot(water_content, aes(x=SOC, y=WHC, color=treatment), size=6) + geom_point() + geom_smooth(method=lm) + scale_color_manual(values=wes_palette(n=3, name="Moonrise2")) + labs(colour = "Treatment") + xlab("Soil organic carbon [%]") + ylab("Water holding capacity [%]")

print(p_whc_carbon_treat)

#ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/regression_whc_carbon_treat.jpeg", width = 20, height = 10, units = "cm")

```
