---
title: "environment_fg"
author: "Katja Kozjek"
output: 
  rmarkdown::html_document:
    code_folding: 'hide'
    toc: true
    toc_float: true
    smart: true
    number_sections: false
    highlight: tango
    self_contained: true
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r wrap-hook, include=FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

#load packages 

```{r setting the environment, message=F, warning=F}

library(car)
library(lmerTest)
library(lmtest)
library(multcomp)
library(ggplot2)
library(wesanderson)
library(emmeans)
library(cowplot)
library(ggpubr)
library(effects)
```

***IMPORTANT: drought treatment, and country are taken as factors***
***SOC as continuous variable*** 

***because 1. field pairs do not always differ in SOC, 2. “low” and “high” present very different values not just among countries, but also within countries and 3. using SOC as a continuous gives us an idea about the actual values that matter, field pairs are accounted in the random structure***

# Water content 

## water content (g water/g soil)

```{r water content, message=FALSE, warning=FALSE}
water_contennt <- read.csv("data/metafile/water_content.csv")

data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      se = sd(x[[col]], na.rm=TRUE) / sqrt(sum(!is.na(x[[col]]))))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
  return(data_sum)
}

water_summary <- data_summary(water_content, varname="WC", 
                 groupnames=c("treatment", "country"))

water_summary$treatment<-factor(water_summary$treatment, levels=c("C", "RC", "R"))
water_summary$country<-factor(water_summary$country, levels=c("SE", "DE", "ES"),labels = c("Sweden", "Germany", "Spain"))

#plot soil water content 
p_water <- ggplot(water_summary, aes(treatment, WC)) +
    geom_errorbar(
    aes(ymin = WC-se, ymax = WC+se, color = treatment),
    position = position_dodge(0.3), width = 0.2
    ) +
    geom_point(aes(color = treatment), position = position_dodge(0.3)) + facet_grid(.~country) + theme_bw() + xlab("Drought treatment") + ylab("Soil water content [%]") + scale_color_manual(values = c("#B6854D", "#8D8680", "#0F0D0E")) + labs(color = "Drought treatment") + theme(axis.text = element_text(size = 10)) + 
  theme(axis.title = element_text(size = 12)) +
  theme(legend.title = element_text(size = 12)) + 
  theme(legend.text = element_text(size = 12)) + 
  theme(strip.text.x = element_text(size = 12))

print(p_water)
```
# Soil carbon content 

## organic carbon 

```{r organic carbon, message=FALSE, warning=FALSE}
metafile <- read.csv("data/metafile/metafile_relaxed.csv, sep=",", row.names=1)

#organic carbon content as continuous 
carbon_summary <- data_summary(metafile, varname="organic.C", 
                 groupnames=c("carbon_intensity", "field_pair", "country"))

carbon_summary$carbon_intensity<-factor(carbon_summary$carbon_intensity, levels=c("L", "H"))
carbon_summary$field_pair<-factor(carbon_summary$field_pair, levels=c("1", "2", "3", "4", "5"))
carbon_summary$country<-factor(carbon_summary$country, levels=c("SE", "DE", "ES"),labels = c("Sweden", "Germany", "Spain"))

#plot % of SOC in high and low
p_organicC <- ggplot(carbon_summary, aes(x=field_pair, y=organic.C, group=carbon_intensity, color=carbon_intensity)) +
    geom_errorbar(
    aes(ymin = organic.C-se, ymax = organic.C+se, color = carbon_intensity),
    position = position_dodge(0.3), width = 0.2
    ) +
    geom_point(aes(color = carbon_intensity), position = position_dodge(0.3)) + facet_grid(.~country) + theme_bw() + xlab("Field site") + ylab("SOC [%]") + scale_color_manual(values = c("#5B1A18", "#D67236")) + labs(color = "SOC category") + theme(axis.text = element_text(size = 20)) + theme(axis.text = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12)) +
  theme(legend.title = element_text(size = 12)) + 
  theme(legend.text = element_text(size = 12)) + 
  theme(strip.text.x = element_text(size = 12))

print(p_organicC) 
```

# Mixed models 

## model 1 
+ https://ademos.people.uic.edu/Chapter18.html

```{r model 1, message=FALSE, warning=FALSE}
#model for WC
#carbon as actual value (continuous), as covariate
#plot gets an ID, from 1-30, explicit nesting -> take it as random effect
#ignore field pair

water_content$Site <- paste0(water_content$country, "_", water_content$field_pair)

mod1 <- lmer(WC~country*treatment + SOC + (1|Site) , data=water_content, REML = TRUE)
#test significance of the model
set.seed(7393
anova_mod1 <- anova(mod1, type = "III")
anova_mod1 
summary(mod1)
coef(summary(mod1))

#Check assumptions of the model
qqnorm(resid(mod1)); qqline(resid(mod1)) 
plot(mod1) 

#Check residuals for each variable (independence and homogeneity)
#plot(resid(mod1)~water_content$country) 
#plot(resid(mod1)~water_content$treatment)
```

### effects mod1

```{r effects mod1, message=FALSE, warning=FALSE}
#effects
ef_mod1 <- effect("country*treatment", mod1)
summary(ef_mod1)

ef_mod1_df <- as.data.frame(ef_mod1)

ggplot(ef_mod1_df, aes(country, fit, color=treatment)) + geom_point() + geom_errorbar(aes(ymin=fit-se, ymax=fit+se), width=0.4) + theme_bw(base_size=12) + facet_grid(~treatment)

e1.lm1 <- predictorEffects(mod1)
plot(e1.lm1)

plot(predictorEffects(mod1))
```

### post-hoc mod1

```{r post hoc mod1, message=FALSE, warning=FALSE}
#check the contrasts
summary(glht(mod1, linfct = mcp(treatment = "Tukey")), test = adjusted("holm")) #contrasts between R and C, RC and R
summary(glht(mod1, linfct = mcp(country = "Tukey")), test = adjusted("holm")) #slight contrast between SE and ES

#post hoc test
#Multiple comparison 
m1.emm_country <- emmeans(mod1, ~ country)
m1.emm.res_country <- pairs(m1.emm_country)
m1.emm.res_country #Results are averaged over the levels of: treatment
#DE-SE, ES-SE

write.csv(m1.emm.res_country, "/Users/Katja/Box/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/emmeans_country.csv")

plot(m1.emm.res_country, comparisons = TRUE)

m1.emm_treat <- emmeans(mod1, ~ treatment)
m1.emm.res_treat <- pairs(m1.emm_treat)
m1.emm.res_treat #Results are averaged over the levels of: country
#C-R

write.csv(m1.emm.res_treat, "/Users/Katja/Box/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/emmeans_treatment.csv")

plot(m1.emm.res_treat, comparisons = TRUE)
```

## each country separetly

```{r each country, message=FALSE, warning=FALSE}
#Sweden
sampledata_SE <- read.csv("data/metafile/sampledata_SE.csv", row.names = 1)

#Germany
sampledata_DE <- read.csv("data/metafile/sampledata_DE.csv", row.names = 1)

#Spain
sampledata_ES <- read.csv("data/metafile/sampledata_SE.csv", row.names = 1)
```

# Regression
http://www.sthda.com/english/wiki/ggplot2-scatter-plots-quick-start-guide-r-software-and-data-visualization

## regression WC-SOC country

```{r simple regression water-SOC country, message=FALSE, warning=FALSE}
p_water_carbon_country <- ggplot(water_content, aes(x=SOC, y=WC, color=country), size=6) + geom_point() + geom_smooth(method=lm) + scale_color_manual(values = c("#9C964A", "#79402E", "#46ACC8")) + labs(colour = "Climatic region") + xlab("SOC [%]") + ylab("Soil water content [%]") + theme_bw() + theme(axis.text = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12)) +
  theme(legend.title = element_text(size = 12)) + 
  theme(legend.text = element_text(size = 12)) 

print(p_water_carbon_country)
```
