---
title: "EZfam_counts"
author: "Katja Kozjek"
date: "17/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r load libraries, include=FALSE}

#load libraries
library(edgeR)
library(RColorBrewer)
library(vegan)
library(phyloseq)
library(devtools)
library(ggplot2)
library(DESeq2)
library(dplyr)
library(tibble)
library(scales)
library(wesanderson)
library(cowplot)

```

# Preprocessing
+ https://ucdavis-bioinformatics-training.github.io/2018-June-RNA-Seq-Workshop/thursday/DE.html 
+ https://informatics.fas.harvard.edu/workshops/HarvardInformatics_DEworkshop_Spring2018.html
+ https://www.bioconductor.org/packages/devel/workflows/vignettes/RNAseq123/inst/doc/limmaWorkflow.html#data-packaging

```{r create DGE object, echo=F}

TDB_EZfam <- readRDS("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/rds/TDB_EZfam.rds")

#original matrix
dim(TDB_EZfam) #309 90
sum(TDB_EZfam) #29241527

#Create DGEList object, non-filtered
EZfam_dgelist <- DGEList(TDB_EZfam)

sum(EZfam_dgelist$counts) #29241527
head(EZfam_dgelist)
class(EZfam_dgelist)

#without 2 spanish samples

TDB_EZfam_remove <- readRDS("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/rds/TDB_EZfam_remove.rds")

dim(TDB_EZfam_remove) #309 88
sum(TDB_EZfam_remove) #29133262

EZfam_dgelist_rm <- DGEList(TDB_EZfam_remove)
sum(EZfam_dgelist_rm$counts)
class(EZfam_dgelist_rm)

```

## Sample information

```{r organise sample information, echo=F}

EZ_names <- read.csv("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/limma/EZ_names.csv", row.names = 1)
str(EZ_names)

#without spanish samples
EZ_names_rm <- EZ_names[-c(58,60),]

all(rownames(EZ_names) == colnames(EZfam_dgelist))
all(rownames(EZ_names_rm) == colnames(EZfam_dgelist_rm))

#for Spanish samples, otherwise remove rm
da_treat <- as.factor(EZ_names_rm$treatment)
da_country <- as.factor(EZ_names_rm$country)
da_SOC <- as.numeric(EZ_names_rm$organic.C)
da_treat_country <- as.factor(EZ_names_rm$treat_country)
da_drought_control <- as.factor(EZ_names_rm$drought_control)
da_pH <- as.numeric(EZ_names_rm$pH)
da_VWC <- as.numeric(EZ_names_rm$VWC)
da_N <- as.numeric(EZ_names_rm$total.N)
da_P <- as.numeric(EZ_names_rm$total.P)
da_reduction <- as.numeric(EZ_names_rm$reduction)

```

## TMM normalization original 
+ all samples

```{r normalising gene expression distributions original, echo=F}

#Calculate normalization factors
#Note: calcNormFactors doesn’t normalize the data, it just calculates normalization factors for use downstream

EZfam_dgelist <- calcNormFactors(EZfam_dgelist,method=c("TMM")) #trimmed mean of M-values
EZfam_dgelist$counts
sum(EZfam_dgelist$counts) 

```

## TMM normalization without ES

```{r normalising gene expression without ES, echo=F}

EZfam_dgelist_rm <- calcNormFactors(EZfam_dgelist_rm,method=c("TMM")) #trimmed mean of M-values
sum(EZfam_dgelist_rm$counts) 
29133262/29241527 #99.6

```

## Remove genes
+ using the FilterByExpr()
+ https://rdrr.io/bioc/edgeR/man/filterByExpr.html
+ use with or without Spanish samples

+ here now without 2 Spanish samples

```{r remove lowly expressed genes, echo=F}

table(rowSums(EZfam_dgelist_rm$counts==0)==90)
#FALSE 309, all genes have counts across samples

#automatic way to filter genes
keep.exprs <- filterByExpr(EZfam_dgelist_rm)

#number of genes left after filtering
EZfam_dgelist_rm <- EZfam_dgelist_rm[keep.exprs,, keep.lib.sizes=FALSE]
dim(EZfam_dgelist_rm) #173 90 with Spanish samples
#175  88 without Spanish s.

sum(EZfam_dgelist$counts) #29166424
29166424/29241527 #we keep 99% reads

#without Spanish s.
sum(EZfam_dgelist_rm$counts) #29063458
29063458/29241527 #we keep 99% reads

write.table(EZfam_dgelist$counts, "C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/TDB_EZfam_filt.counts.tsv",row.names = TRUE, col.names = TRUE, sep = "\t" )

write.table(EZfam_dgelist_rm$counts, "C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/TDB_EZfam_rem_filt.counts.tsv",row.names = TRUE, col.names = TRUE, sep = "\t" )

```
***175 genes left after filtering from the original matrix, without 2 samples.***

```{r PCA without ES, echo=F}

#remove spanish samples
#PCA with filterByExpr

tdb_fam <- read.csv("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/TDB_EZfam_filt.counts.tsv",sep = "\t", header = T, row.names = 1)

tdb_fam <- tdb_fam[,-c(58,60)]

metafile_pca <- read.csv("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/metafile_relaxed.csv", sep=",")
metafile_pca_remove <- metafile_pca[-c(58,60),]

dds_fam <- DESeqDataSetFromMatrix(countData = tdb_fam,
                              colData = metafile_pca_remove,
                              design= ~ country + treatment)

vsd_fam <- varianceStabilizingTransformation(dds_fam)

p1 <- plotPCA(vsd_fam, "country")
p2 <- plotPCA(vsd_fam, "treatment")
plot_grid(p1, p2)

#batch correction
vsd_fam_remCOUNTRY <- vsd_fam
assay(vsd_fam_remCOUNTRY) <- limma::removeBatchEffect(assay(vsd_fam_remCOUNTRY), vsd_fam_remCOUNTRY$country)

p1 <- plotPCA(vsd_fam_remCOUNTRY, "country")
p2 <- plotPCA(vsd_fam_remCOUNTRY, "treatment")
plot_grid(p1, p2)

```

# Differential abundance analysis
## MDS plot

```{r MDS plot original filtered, echo = FALSE}

#Datasets where samples do not cluster by experimental group may show little or no evidence of differential expression in the downstream analysis.

#country
col.country <- da_country
levels(col.country) <-  brewer.pal(nlevels(col.country), "Set1")
col.country  <- as.character(col.country)

plotMDS(EZfam_dgelist_rm, labels=da_country, col=col.country) #DE + SE cluster together

#treatment
col.treatment <- da_treat
levels(col.treatment) <-  brewer.pal(nlevels(col.treatment), "Set1")
col.treatment <- as.character(col.treatment)

plotMDS(EZfam_dgelist_rm, labels=da_treat, col=col.treatment, dim=c(1,2))
plotMDS(EZfam_dgelist_rm, labels=da_treat, col=col.treatment, dim=c(3,4)) #no clustering according to the treatment

```

## Design matrix

```{r create design matrix, echo=F}

#treatment
design_treat <- model.matrix(~0+da_treat) #if you remove 0 then it takes treatment C as intercept and serve as control
colnames(design_treat) <- gsub("treatment", "", colnames(design_treat))
colnames(design_treat)

#country 
design_country <- model.matrix(~0+da_country)
colnames(design_country) <- gsub("country", "", colnames(design_country))
colnames(design_country)

#If we had other covariates to control for, the variable of interest appears in the first position, here we control for the effect of country or the effect of treatment
mm1 <- model.matrix(~0 + da_country + da_treat) #treatment as batch 
mm2 <- model.matrix(~0 + da_treat + da_country) #country as batch

```

## Overall DA model 

### treatment as batch 

```{r model 1 overall DA, echo=F}

#voom model1: mean-variance trend
v_mm1 <- voom(EZfam_dgelist_rm, mm1, plot=TRUE)

#final model:mean-variance trend
vfit_mm1 <- lmFit(v_mm1, mm1)
head(coef(vfit_mm1))

contr_mm1 <- makeContrasts(da_countryDE - da_countryES, da_countryDE - da_countrySE, da_countryES - da_countrySE, levels = colnames(coef(vfit_mm1)))

vfit_mm1 <- contrasts.fit(vfit_mm1, contr_mm1) #Estimate contrast for each gene
efit_mm1 <- eBayes(vfit_mm1, robust = TRUE) #Empirical Bayes smoothing of standard errors

summary(decideTests(efit_mm1, adjust.method="BH",p.value = 0.05))

topTable(efit_mm1, coef=1, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")
topTable(efit_mm1, coef=2, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")
topTable(efit_mm1, coef=3, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")

```

### country as batch

```{r model 2 overall DA, echo=F}

#voom model2: mean-variance trend
v_mm2 <- voom(EZfam_dgelist_rm, mm2, plot=TRUE)

#final model:mean-variance trend
vfit_mm2 <- lmFit(v_mm2, mm2)
head(coef(vfit_mm2))

contr_mm2 <- makeContrasts(da_treatR - da_treatC, da_treatRC - da_treatC, da_treatR - da_treatRC, levels = colnames(coef(vfit_mm2)))

vfit_mm2 <- contrasts.fit(vfit_mm2, contr_mm2) #Estimate contrast for each gene
efit_mm2 <- eBayes(vfit_mm2, robust = TRUE) #Empirical Bayes smoothing of standard errors

summary(decideTests(efit_mm2, adjust.method="BH",p.value = 0.05))

```

## Contrasts 
+ between drought treatments
+ between locations

```{r contrasts, echo=F}

#treatment
contr.matrix_treat <- makeContrasts(
   CvsRC = da_treatC - da_treatRC, 
   RCvsR = da_treatRC - da_treatR, 
   CvsR = da_treatC - da_treatR, 
   levels = colnames(design_treat))

#location
contr.matrix_country <- makeContrasts(
   SEvsES = da_SE - da_ES,
   SEvsDE = da_SE - da_DE, 
   DEvsES = da_DE - da_ES, 
   DESEvsES = (da_DE + da_SE) - da_ES,
   levels = colnames(design_country))

```

## Model for treatment

```{r model treatment, echo=F}

#voom model: mean-variance trend
v_treat <- voom(EZfam_dgelist_rm, design_treat, plot=TRUE)

#final model:mean-variance trend
vfit_treat <- lmFit(v_treat, design_treat)
head(coef(vfit_treat))

vfit_treat <- contrasts.fit(vfit_treat, contr.matrix_treat) #Estimate contrast for each gene
efit_treat <- eBayes(vfit_treat, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_treat, main="Final model: Mean-variance trend")

```
***Means (x-axis) and variances (y-axis) of each gene are plotted***
***first plot: before voom is applied to the data***
***second plot: the trend is removed after voom precision weights are applied to the data***

### DA genes treatment

```{r number of DA genes in treatment, echo=F}

#Examining the number of DA genes
summary(decideTests(efit_treat, adjust.method="BH",p.value = 0.05))

topTable(efit_treat, coef = "RCvsR", adjust.method = "BH", p.value = 0.05)
topTable(efit_treat, coef = "CvsRC", adjust.method = "BH", p.value = 0.05)
topTable(efit_treat, coef = "CvsR", adjust.method = "BH", p.value = 0.05)

```
***No contrast: 175 up-regulated in all conditions***
***Contrast:175 non-significant***

## Model for country

```{r model country, echo=F}

#voom model: mean-variance trend
v_country <- voom(EZfam_dgelist_rm, design_country, plot=TRUE)

#final model:mean-variance trend
vfit_country <- lmFit(v_country, design_country)
head(coef(vfit_country))

vfit_country <- contrasts.fit(vfit_country, contr.matrix_country) #Estimate contrast for each gene
efit_country <- eBayes(vfit_country, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_country, main="Final model: Mean-variance trend")

```

### DA genes country

```{r number of DA genes in country, echo=F}

#the number of significantly up- and down-regulated genes
#based on an adjusted p-value cutoff that is set at 5% by default
#BH is equivalent with fdr

#Examining the number of DE genes
summaryDA_country <- summary(decideTests(efit_country,adjust.method="BH",p.value = 0.05))
summaryDA_country
      #       SEvsES SEvsDE DEvsES
#Down       62     38     68
#NotSig     48     96     42
#Up         65     41     65

```
***62 genes are found to be down-regulated in SE relative to ES, 65 are up-regulated - a total of 127 DA genes***
***79 DA genes are found between SE and DE (38 down, 41 up)***
***133 DA genes are found between DE and ES (68 down, 65 up)***
***No contrast:175 up-regulated***

```{r top DA location, echo=F}

#(Permitted synonyms are "M" for "logFC", "A" or "Amean" for "AveExpr", "T" for "t" and "p" for "P".) 
#genes from smallest to largest adjusted p-value

#What genes are most differentially abundant?
top_SEvsES <- topTable(efit_country, coef=1, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")

head(top_SEvsES, 5)

top_SEvsDE <- topTable(efit_country, coef=2, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")

head(top_SEvsDE, 5)

top_DEvsES <- topTable(efit_country, coef=3, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")

head(top_DEvsES, 5)

#(Germany + Sweden) - Spain
top_DESEvsES <- topTable(efit_country, coef=4, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")

head(top_DESEvsES, 10)

#write table
#top_country$enzyme_family <- rownames(top_country)
#top_country <- top_country[,c("enzyme_family", names(top_country)[1:8])]
#top_country 
#write.table(top_country, file = "contrasts_country.txt", row.names = F, sep = "\t", quote = F)

```

# Correlation of gene abundance with a continuous variable? 
+ how different genes are abundant along with a continuous variable 
+ matrix without 2 Spanish samples used

## SOC 

```{r organic C, echo=F}

model_SOC <- model.matrix(~da_SOC)
head(model_SOC)

v_SOC <- voom(EZfam_dgelist_rm, model_SOC, plot = T)

fit_SOC <- lmFit(v_SOC, model_SOC)
tmp_SOC <- contrasts.fit(fit_SOC, coef = 2) # test "SOC" coefficient

tmp_SOC <- eBayes(tmp_SOC)
summary(decideTests(tmp_SOC,adjust.method="BH", p.value=0.05))

top.table_SOC <- topTable(tmp_SOC, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")
head(top.table_SOC, 5)

```

```{r plot SOC genes, echo=F}

CE1 <- v_SOC$E["CE1",]
plot(CE1 ~ da_SOC) #average expression/abundance is plotted
intercept <- coef(fit_SOC)["CE1", "(Intercept)"]
slope <- coef(fit_SOC)["CE1", "da_SOC"]
abline(a = intercept, b = slope)

sampleID <- c("DE","DE" ,"DE" ,"DE" ,"DE" ,"DE" ,"DE" ,"DE" ,"DE" ,"DE" ,"DE" ,"DE" ,"DE" ,"DE" ,"DE" ,"DE" ,"DE","DE","DE" ,"DE" ,"DE" ,"DE", "DE" ,"DE" ,"DE" ,"DE" ,"DE", "DE" ,"DE" ,"DE" ,"ES" , "ES"  ,"ES"  ,"ES"  ,"ES"  ,"ES" ,"ES"  ,"ES"  ,"ES" , "ES" ,"ES" ,"ES" ,"ES" ,"ES" ,"ES","ES", "ES" ,"ES" ,"ES" ,"ES" ,"ES" ,"ES" ,"ES" ,"ES","ES" ,"ES" ,"ES", "ES" ,"SE"  ,"SE"  ,"SE" ,"SE"  ,"SE"  ,"SE"  ,"SE"  ,"SE"  ,"SE" , "SE" ,"SE" ,"SE","SE" ,"SE" ,"SE", "SE" ,"SE" ,"SE" ,"SE" ,"SE" ,"SE","SE" ,"SE" ,"SE" ,"SE" ,"SE" ,"SE","SE" ,"SE","SE")

CE1 <- as.data.frame(CE1)
CE1bind <- cbind(CE1, sampleID)

plot_CE1 <- ggplot(CE1bind) +
  geom_point(aes(da_SOC, CE1, color = factor(sampleID)), 
             position = 'jitter') +  geom_abline(slope = slope, intercept = intercept) + scale_x_continuous("SOC") + 
   labs(colour = "Country") + theme_bw() +  scale_color_manual(values = c("#9C964A", "#B6854D","#46ACC8"))

CE4 <- v_SOC$E["CE4",]
plot(CE4 ~ da_SOC) #average expression/abundance is plotted
intercept_CE4 <- coef(fit_SOC)["CE4", "(Intercept)"]
slope_CE4 <- coef(fit_SOC)["CE4", "da_SOC"]
abline(a = intercept, b = slope)

CE4 <- as.data.frame(CE4)
CE4bind <- cbind(CE4, sampleID)

plot_CE4 <- ggplot(CE4bind) +
  geom_point(aes(da_SOC, CE4, color = factor(sampleID)), 
             position = 'jitter') +  geom_abline(slope = slope_CE4, intercept = intercept_CE4) + scale_x_continuous("SOC") + 
   labs(colour = "Country") + theme_bw() +  scale_color_manual(values = c("#9C964A", "#B6854D","#46ACC8"))


GT51 <- v_SOC$E["GT51",]
plot(GT51 ~ da_SOC) #average expression/abundance is plotted
intercept_GT51 <- coef(fit_SOC)["GT51", "(Intercept)"]
slope_GT51 <- coef(fit_SOC)["GT51", "da_SOC"]
abline(a = intercept, b = slope)

GT51 <- as.data.frame(GT51)
GT51bind <- cbind(GT51, sampleID)

plot_GT51 <- ggplot(GT51bind) +
  geom_point(aes(da_SOC, GT51, color = factor(sampleID)), 
             position = 'jitter') +  geom_abline(slope = slope_GT51, intercept = intercept_GT51) + scale_x_continuous("SOC") + 
   labs(colour = "Country") + theme_bw() +  scale_color_manual(values = c("#9C964A", "#B6854D","#46ACC8"))


CBM50 <- v_SOC$E["CBM50",]
plot(CBM50 ~ da_SOC) #average expression/abundance is plotted
intercept_CBM50 <- coef(fit_SOC)["CBM50", "(Intercept)"]
slope_CBM50 <- coef(fit_SOC)["CBM50", "da_SOC"]
abline(a = intercept, b = slope)

CBM50 <- as.data.frame(CBM50)
CBM50bind <- cbind(CBM50, sampleID)

plot_CBM50 <- ggplot(CBM50bind) +
  geom_point(aes(da_SOC, CBM50, color = factor(sampleID)), 
             position = 'jitter') +  geom_abline(slope = slope_CBM50, intercept = intercept_CBM50) + scale_x_continuous("SOC") + 
   labs(colour = "Country") + theme_bw() +  scale_color_manual(values = c("#9C964A", "#B6854D","#46ACC8"))

plot_grid(plot_CE1, plot_CE4, plot_GT51, plot_CBM50)

ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/SOC_genes.jpeg", width = 20, height = 15, units = "cm")

```

## pH

```{r pH, echo=F}

model_pH <- model.matrix(~da_pH)
head(model_pH)

v_pH <- voom(EZfam_dgelist_rm, model_pH, plot = T)

fit_pH <- lmFit(v_pH, model_pH)
tmp_pH <- contrasts.fit(fit_pH, coef = 2) # test "pH" coefficient

tmp_pH <- eBayes(tmp_pH)
summary(decideTests(tmp_pH, adjust.method="BH", p.value=0.05))

top.table_pH <- topTable(tmp_pH,number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")
head(top.table_pH, 5)

CBM2 <- v_pH$E["CBM2",]
plot(CBM2 ~ da_pH)
intercept_pH_CBM2 <- coef(fit_pH)["CBM2", "(Intercept)"]
slope_pH_CBM2 <- coef(fit_pH)["CBM2", "da_pH"]
abline(a = intercept_pH_CBM2, b = slope_pH_CBM2)

CBM2 <- as.data.frame(CBM2)
CBM2bind <- cbind(CBM2, sampleID)

plot_CBM2 <- ggplot(CBM2bind) +
  geom_point(aes(da_pH, CBM2, color = factor(sampleID)), 
             position = 'jitter') +  geom_abline(slope = slope_pH_CBM2, intercept = intercept_pH_CBM2) + scale_x_continuous("pH") + 
   labs(colour = "Country") + theme_bw() +  scale_color_manual(values = c("#9C964A", "#B6854D","#46ACC8"))


PL11 <- v_pH$E["PL11",]
plot(PL11 ~ da_pH)
intercept_pH_PL11 <- coef(fit_pH)["PL11", "(Intercept)"]
slope_pH_PL11 <- coef(fit_pH)["PL11", "da_pH"]
abline(a = intercept_pH_PL11, b = slope_pH_PL11)

PL11 <- as.data.frame(PL11)
PL11bind <- cbind(PL11, sampleID)

plot_PL11 <- ggplot(PL11bind) +
  geom_point(aes(da_pH, PL11, color = factor(sampleID)), 
             position = 'jitter') +  geom_abline(slope = slope_pH_PL11, intercept = intercept_pH_PL11) + scale_x_continuous("pH") + 
   labs(colour = "Country") + theme_bw() +  scale_color_manual(values = c("#9C964A", "#B6854D","#46ACC8"))

plot_grid(plot_CBM2, plot_PL11)
ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/ph_genes.jpeg", width = 20, height = 10, units = "cm")

```

## Soil moisture (=VWC)

```{r soil moisture, echo=F}

model_moisture <- model.matrix(~da_water_content)
head(model_moisture)

v_moisture <- voom(EZfam_dgelist_rm, model_moisture, plot = T)

fit_moisture <- lmFit(v_moisture, model_moisture)
tmp_moisture <- contrasts.fit(fit_moisture, coef = 2) # test "WC/WHC" coefficient

tmp_moisture <- eBayes(tmp_moisture)
summary(decideTests(tmp_moisture, adjust.method="BH", p.value=0.05))

top.table_moisture <- topTable(tmp_moisture, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")
head(top.table_moisture, 5)

GH67 <- v_moisture$E["GH67",]
plot(GH67 ~ da_water_content)
intercept_WC <- coef(fit_moisture)["GH67", "(Intercept)"]
slope_WC <- coef(fit_moisture)["GH67", "da_water_content"]
abline(a = intercept_WC, b = slope_WC)

plot_WC <- data.frame(cbind(intercept_WC, slope_WC, GH67, da_water_content))
ggplot(data=plot_WC, aes(x=da_water_content, y=GH67)) + geom_abline(data=plot_WC,aes(slope=slope_WC,intercept=intercept_WC))

```

# Separate countries
## Sweden

### Prep

```{r Sweden prep, echo=F}

dim(EZfam_SE) #309 90

#Create DGEList object, non-filtered
EZfamSE_dgelist <- DGEList(EZfam_SE)

head(EZfamSE_dgelist)
dim(EZfamSE_dgelist)

#original
table(rowSums(EZfamSE_dgelist$counts==0)==30)
#FALSE 291, TRUE 18, 18 genes in this dataset have zero counts across all 30 samples

#automatic way to filter genes
keep.exprsSE <- filterByExpr(EZfamSE_dgelist)

#number of genes left after filtering
EZfamSE_dgelist <- EZfamSE_dgelist[keep.exprsSE,, keep.lib.sizes=FALSE]
dim(EZfamSE_dgelist) #174 30
sum(EZfamSE_dgelist$counts) #8169176

``` 


### Sample info 

```{r Sweden metafile, echo=F}

EZfamSE_names <- EZfam_names[c(61:90),]
all(rownames(EZfamSE_names) == colnames(EZfamSE_dgelist))

treatmentSE <- as.factor(EZfamSE_names$treatment)
SOC_SE <- as.numeric(EZfamSE_names$organic.C)
drought_control_SE <- as.factor(EZfamSE_names$drought_control)
pH_SE <- as.numeric(EZfamSE_names$pH)
water_content_SE <- as.numeric(EZfamSE_names$WC)
nitrogen_SE <- as.numeric(EZfamSE_names$total.N)
phosphorus_SE <- as.numeric(EZfamSE_names$total.P)
fieldSE <- as.numeric(EZfamSE_names$field_pair)

``` 

### Normalization

```{r SE normalising gene expression distributions, echo=F}

EZfamSE_dgelist <- calcNormFactors(EZfamSE_dgelist,method =c("TMM")) #trimmed mean of M-values

```

### MDS plot

```{r plot MDS SE, echo=F}

#treatment
col.treatmentSE <- treatmentSE 
levels(col.treatmentSE) <-  brewer.pal(nlevels(col.treatmentSE), "Set1")
col.treatmentSE <- as.character(col.treatmentSE)

plotMDS(EZfamSE_dgelist, labels=treatmentSE, col=col.treatmentSE, dim=c(1,2))

#field pair
col.fieldSE <- fieldSE 
levels(col.fieldSE) <-  brewer.pal(nlevels(col.fieldSE), "Set1")
col.fieldSE <- as.character(col.fieldSE)

plotMDS(EZfamSE_dgelist, labels=fieldSE, col=col.fieldSE, dim=c(1,2))

```
***From the plot it seems there is no DA between treatments***

### SE model for treatment

```{r Sweden model, echo=F}

design_SEtreat <- model.matrix(~0+treatmentSE) #if you remove 0 then it takes treatment C as intercept and serve as control
colnames(design_SEtreat)

#voom model: mean-variance trend
v_SEtreat <- voom(EZfamSE_dgelist, design_SEtreat, plot=TRUE)

#final model:mean-variance trend
vfit_SEtreat <- lmFit(v_SEtreat, design_SEtreat)
head(coef(vfit_SEtreat))

#contrast
contr_SEtreat <- makeContrasts(treatmentSERC - treatmentSER, treatmentSEC - treatmentSER, treatmentSEC - treatmentSERC, levels = colnames(coef(vfit_SEtreat)))
contr_SEtreat

vfit_SEtreat <- contrasts.fit(vfit_SEtreat, contr_SEtreat)
efit_SEtreat <- eBayes(vfit_SEtreat, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_SEtreat, main="Final model: Mean-variance trend")

summary(decideTests(efit_SEtreat, adjust.method="BH",p.value = 0.05)) #all non-significant

```
***Contrast:174 non-significant***

```{r Sweden only R, echo=F}

##R vs RC+C

design_SE_drought <- model.matrix(~drought_control_SE)
colnames(design_SE_drought)

#voom model: mean-variance trend
v_SEdrought <- voom(EZfamSE_dgelist, design_SE_drought, plot=TRUE)

#final model:mean-variance trend
vfit_SEdrought <- lmFit(v_SEdrought, design_SE_drought)
head(coef(vfit_SEdrought))

efit_SEdrought <- eBayes(vfit_SEdrought, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_SEdrought, main="Final model: Mean-variance trend")

summary(decideTests(efit_SEdrought, adjust.method="BH", p.value=0.05))

```

## Germany

### Prep

```{r Germany prep}

dim(EZfam_DE) #309 90

#Create DGEList object, non-filtered
EZfamDE_dgelist <- DGEList(EZfam_DE)

head(EZfamDE_dgelist)
dim(EZfamDE_dgelist)

#original
table(rowSums(EZfamDE_dgelist$counts==0)==30)
#FALSE 278, TRUE 31, 31 genes in this dataset have zero counts across all 30 samples

#automatic way to filter genes
keep.exprsDE <- filterByExpr(EZfamDE_dgelist)

#number of genes left after filtering
EZfamDE_dgelist <- EZfamDE_dgelist[keep.exprsDE,, keep.lib.sizes=FALSE]
dim(EZfamDE_dgelist) #172 30
sum(EZfamDE_dgelist$counts) #7242010

``` 

### Sample info 

```{r Germany metafile}

EZfamDE_names <- EZfam_names[c(1:30),]
all(rownames(EZfamDE_names) == colnames(EZfamDE_dgelist))

treatmentDE <- as.factor(EZfamDE_names$treatment)
SOC_DE <- as.numeric(EZfamDE_names$organic.C)
drought_control_DE <- as.factor(EZfamDE_names$drought_control)
pH_DE <- as.numeric(EZfamDE_names$pH)
water_content_DE <- as.numeric(EZfamDE_names$WC)
nitrogen_DE <- as.numeric(EZfamDE_names$total.N)
phosphorus_DE <- as.numeric(EZfamDE_names$total.P)
fieldDE <- as.numeric(EZfamDE_names$field_pair)

``` 

### Normalization

```{r DE normalising gene expression distributions}

EZfamDE_dgelist <- calcNormFactors(EZfamDE_dgelist,method =c("TMM")) #trimmed mean of M-values

```

### MDS plot

```{r plot MDS DE}

#treatment
col.treatmentDE <- treatmentDE 
levels(col.treatmentDE) <-  brewer.pal(nlevels(col.treatmentDE), "Set1")
col.treatmentDE <- as.character(col.treatmentDE)

plotMDS(EZfamDE_dgelist, labels=treatmentDE, col=col.treatmentDE, dim=c(1,2))

#field pair
col.fieldDE <- fieldDE 
levels(col.fieldDE) <-  brewer.pal(nlevels(col.fieldDE), "Set1")
col.fieldDE <- as.character(col.fieldDE)

plotMDS(EZfamDE_dgelist, labels=fieldDE, col=col.fieldDE, dim=c(1,2))

```
***From the plot is seems there is no DA between treatments***

### DE model for treatment

```{r Germany model}

design_DEtreat <- model.matrix(~0+treatmentDE)
colnames(design_DEtreat)

#voom model: mean-variance trend
v_DEtreat <- voom(EZfamDE_dgelist, design_DEtreat, plot=TRUE)

#final model:mean-variance trend
vfit_DEtreat <- lmFit(v_DEtreat, design_DEtreat)
head(coef(vfit_DEtreat))

#contrast
contr_DEtreat <- makeContrasts(treatmentDERC - treatmentDER, treatmentDEC - treatmentDER, treatmentDEC - treatmentDERC, levels = colnames(coef(vfit_DEtreat)))
contr_DEtreat

vfit_DEtreat <- contrasts.fit(vfit_DEtreat, contr_DEtreat)
efit_DEtreat <- eBayes(vfit_DEtreat, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_DEtreat, main="Final model: Mean-variance trend")

summary(decideTests(efit_DEtreat, adjust.method = "BH", p.value = 0.05))

```
***Contrast:172 non-significant***

```{r Germany only R}

#What genes are most differentially expressed/abundant?
##R vs RC+C

design_DE_drought <- model.matrix(~drought_control_DE) 
colnames(design_DE_drought)

#voom model: mean-variance trend
v_DEdrought <- voom(EZfamDE_dgelist, design_DE_drought, plot=TRUE)

#final model:mean-variance trend
vfit_DEdrought <- lmFit(v_DEdrought, design_DE_drought)
head(coef(vfit_DEdrought))

efit_DEdrought <- eBayes(vfit_DEdrought, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_DEdrought, main="Final model: Mean-variance trend")

summary(decideTests(efit_DEdrought, adjust.method="BH", p.value=0.05))

```
***172 up-regulated***

```{r Germany fields}

design_DE_field <- model.matrix(~fieldDE) 
colnames(design_DE_field)

#voom model: mean-variance trend
v_DEfield <- voom(EZfamDE_dgelist, design_DE_field, plot=TRUE)

#final model:mean-variance trend
vfit_DEfield <- lmFit(v_DEfield, design_DE_field)
head(coef(vfit_DEfield))

efit_DEfield <- eBayes(vfit_DEfield, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_DEfield, main="Final model: Mean-variance trend")

summary(decideTests(efit_DEfield, adjust.method="BH", p.value=0.05))
topTable(efit_DEfield, coef=2, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")

```

## Spain

### Prep

```{r Spain prep}

dim(EZfam_ES_remove) #309  28

#Create DGEList object, non-filtered
EZfamES_dgelist <- DGEList(EZfam_ES_remove)

head(EZfamES_dgelist)
dim(EZfamES_dgelist)

#original
table(rowSums(EZfamES_dgelist$counts==0)==28)
#FALSE 284, TRUE 25, 25 genes in this dataset have zero counts across all 28 samples

#automatic way to filter genes
keep.exprsES <- filterByExpr(EZfamES_dgelist)

#number of genes left after filtering
EZfamES_dgelist <- EZfamES_dgelist[keep.exprsES,, keep.lib.sizes=FALSE]
dim(EZfamES_dgelist) #176  28

``` 

### Sample info 

```{r Spain metafile}

EZfamES_names <- EZfam_names[c(31:60),]
EZfamES_names <- EZfamES_names[-c(28,30),]
all(rownames(EZfamES_names) == colnames(EZfamES_dgelist))

treatmentES <- as.factor(EZfamES_names$treatment)
SOC_ES <- as.numeric(EZfamES_names$organic.C)
drought_control_ES <- as.factor(EZfamES_names$drought_control)
pH_ES <- as.numeric(EZfamES_names$pH)
water_content_ES <- as.numeric(EZfamES_names$WC)
nitrogen_ES <- as.numeric(EZfamES_names$total.N)
phosphorus_ES <- as.numeric(EZfamES_names$total.P)
fieldES <- as.numeric(EZfamES_names$field_pair)

``` 

### Normalization

```{r ES normalising gene expression distributions}

EZfamES_dgelist <- calcNormFactors(EZfamES_dgelist,method =c("TMM")) #trimmed mean of M-values

```

### MDS plot

```{r plot MDS ES}

#treatment
col.treatmentES <- treatmentES 
levels(col.treatmentES) <-  brewer.pal(nlevels(col.treatmentES), "Set1")
col.treatmentES <- as.character(col.treatmentES)

plotMDS(EZfamES_dgelist, labels=treatmentES, col=col.treatmentES, dim=c(1,2))

#field pair
col.fieldES <- fieldES 
levels(col.fieldES) <-  brewer.pal(nlevels(col.fieldES), "Set1")
col.fieldES <- as.character(col.fieldES)

plotMDS(EZfamES_dgelist, labels=fieldES, col=col.fieldES, dim=c(1,2))

```
***From the plot is seems there is no DA between treatments***

### ES model for treatment

```{r Spain model}

design_EStreat <- model.matrix(~0+treatmentES)
colnames(design_EStreat)

#voom model: mean-variance trend
v_EStreat <- voom(EZfamES_dgelist, design_EStreat, plot=TRUE)

#final model:mean-variance trend
vfit_EStreat <- lmFit(v_EStreat, design_EStreat)
head(coef(vfit_EStreat))

#contrast
contr_EStreat <- makeContrasts(treatmentESRC - treatmentESR, treatmentESRC- treatmentESC, treatmentESC -treatmentESR, levels = colnames(coef(vfit_EStreat)))
contr_EStreat

vfit_EStreat <- contrasts.fit(vfit_EStreat, contr_EStreat)
efit_EStreat <- eBayes(vfit_EStreat, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_EStreat, main="Final model: Mean-variance trend")

summary(decideTests(efit_EStreat, adjust.method = "BH", p.value = 0.05))

```
***Contrast:176 non-significant***

```{r Spain top genes}

##R vs RC+C

design_ES_drought <- model.matrix(~drought_control_ES) #if you remove 0 then it takes traetment C as intercept and serve as control
colnames(design_ES_drought)

#voom model: mean-variance trend
v_ESdrought <- voom(EZfamES_dgelist, design_ES_drought, plot=TRUE)

#final model:mean-variance trend
vfit_ESdrought <- lmFit(v_ESdrought, design_ES_drought)
head(coef(vfit_ESdrought))

efit_ESdrought <- eBayes(vfit_ESdrought, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_ESdrought, main="Final model: Mean-variance trend")

summary(decideTests(efit_ESdrought, adjust.method = "BH", p.value = 0.05))

```

```{r Spain fields}

design_ES_field <- model.matrix(~fieldES) 
colnames(design_ES_field)

#voom model: mean-variance trend
v_ESfield <- voom(EZfamES_dgelist, design_ES_field, plot=TRUE)

#final model:mean-variance trend
vfit_ESfield <- lmFit(v_ESfield, design_ES_field)
head(coef(vfit_ESfield))

efit_ESfield <- eBayes(vfit_ESfield, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_ESfield, main="Final model: Mean-variance trend")

summary(decideTests(efit_ESfield, adjust.method="BH", p.value=0.05))
topTable(efit_ESfield, coef=2, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")

```

# TDB EZacc

## Prepare data

```{r TDB EZacc prep}

#original 
dim(TDB_EZacc_relaxed) #55109 90
sum(TDB_EZacc_relaxed) #29241527

#reorder
order_TDB_EZacc <- match(rownames(sampledata), colnames(TDB_EZacc_relaxed))
order_TDB_EZacc  <- TDB_EZacc_relaxed[,order_TDB_EZacc]
head(order_TDB_EZacc)
all(rownames(sampledata) == colnames(order_TDB_EZacc))

#remove Spanish s.
order_TDB_EZacc <- order_TDB_EZacc[,-c(58,60)]

#create dgelist with original matrix EZacc
EZacc_dgelist <- DGEList(order_TDB_EZacc)
dim(EZacc_dgelist) #55109    88

EZacc_dgelist <- calcNormFactors(EZacc_dgelist,method =c("TMM")) #trimmed mean of M-values

table(rowSums(EZacc_dgelist$counts==0)==88)
#FALSE 55080, TRUE 29

#automatic way to filter genes
keep.exprs_EZacc <- filterByExpr(EZacc_dgelist)

#number of genes left after filtering
EZacc_dgelist <- EZacc_dgelist[keep.exprs_EZacc,, keep.lib.sizes=FALSE]
dim(EZacc_dgelist) #1566   88
sum(EZacc_dgelist$counts) #18509701

plotMDS(EZacc_dgelist)
#extract matrix with counts for EZacc
EZacc_counts <- EZacc_dgelist$counts

#write.table(EZacc_dgelist$counts, "C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/TDB_EZacc_rem_filt.counts.tsv",row.names = TRUE, col.names = TRUE, sep = "\t" )

```

## Model EZacc
### treatment 

```{r TDB EZacc model treatment}

#treatment
design_treat_EZacc <- model.matrix(~0+da_treatment)
colnames(design_treat_EZacc)

#voom model: mean-variance trend
v_treat_EZacc <- voom(EZacc_dgelist, design_treat_EZacc, plot=TRUE)

#final model:mean-variance trend
vfit_treat_EZacc <- lmFit(v_treat_EZacc, design_treat_EZacc)

contr.EZacc_treat <- makeContrasts(
   CvsRC = da_treatmentC - da_treatmentRC, 
   RCvsR = da_treatmentRC - da_treatmentR, 
   CvsR = da_treatmentC - da_treatmentR, 
   levels = colnames(design_treat_EZacc))

vfit_treat_EZacc <- contrasts.fit(vfit_treat_EZacc, contr.EZacc_treat)
efit_treat_EZacc <- eBayes(vfit_treat_EZacc, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_treat_EZacc, main="Final model: Mean-variance trend")

summary(decideTests(efit_treat_EZacc, adjust.method = "BH", p.value = 0.05))

```

### location

```{r TDB EZacc model country }

#country
design_country_EZacc <- model.matrix(~0+da_country)
colnames(design_country_EZacc) <- gsub("country", "", colnames(design_country_EZacc))
colnames(design_country_EZacc)

#voom model: mean-variance trend
v_country_EZacc <- voom(EZacc_dgelist, design_country_EZacc, plot=TRUE)

#final model:mean-variance trend
vfit_country_EZacc <- lmFit(v_country_EZacc, design_country_EZacc)

contr.EZacc_country <- makeContrasts(
   SEvsES = da_SE - da_ES,
   SEvsDE = da_SE - da_DE, 
   DEvsES = da_DE - da_ES, 
   levels = colnames(design_country_EZacc))

vfit_country_EZacc <- contrasts.fit(vfit_country_EZacc, contr.EZacc_country)
efit_country_EZacc <- eBayes(vfit_country_EZacc, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_country_EZacc, main="Final model: Mean-variance trend")

summary(decideTests(efit_country_EZacc, adjust.method = "BH", p.value = 0.05))

#         SEvsES SEvsDE DEvsES
#Down      623    298    658
#NotSig    443    933    429
#Up        498    333    477

#What genes are most differentially abundant?
top_EZAcc_SEvsES <- topTable(efit_country_EZacc, coef=1, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")
head(top_EZAcc_SEvsES, 5)

top_EZAcc_SEvsDE <- topTable(efit_country_EZacc, coef=2, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")
head(top_EZAcc_SEvsDE, 5)

top_EZAcc_DEvsES <- topTable(efit_country_EZacc, coef=3, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")
head(top_EZAcc_DEvsES, 5)

```

### overall

```{r overall TDB EZacc}

mm1_EZacc <- model.matrix(~0 + da_country + da_treatment) #treatment as batch 
mm2_EZacc <- model.matrix(~0 + da_treatment + da_country) #country as batch

```

```{r model 1 overall EZacc}

#voom model1: mean-variance trend
v_mm1_EZacc <- voom(EZacc_dgelist, mm1_EZacc, plot=TRUE)

#final model:mean-variance trend
vfit_mm1_EZacc <- lmFit(v_mm1_EZacc, mm1_EZacc)
head(coef(vfit_mm1_EZacc))

contr_mm1_EZacc <- makeContrasts(da_countryDE - da_countryES, da_countryDE - da_countrySE, da_countryES - da_countrySE, levels = colnames(coef(vfit_mm1_EZacc)))

vfit_mm1_EZacc <- contrasts.fit(vfit_mm1_EZacc, contr_mm1_EZacc) #Estimate contrast for each gene
efit_mm1_EZacc <- eBayes(vfit_mm1_EZacc, robust = TRUE) #Empirical Bayes smoothing of standard errors

summary(decideTests(efit_mm1_EZacc, adjust.method="BH",p.value = 0.05))

topTable(efit_mm1_EZacc, coef=1, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")
topTable(efit_mm1_EZacc, coef=2, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")
topTable(efit_mm1_EZacc, coef=3, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")

```

```{r model 2 overall EZacc}

#voom model2: mean-variance trend
v_mm2_EZacc <- voom(EZacc_dgelist, mm2_EZacc, plot=TRUE)

#final model:mean-variance trend
vfit_mm2_EZacc <- lmFit(v_mm2_EZacc, mm2_EZacc)
head(coef(vfit_mm2_EZacc))

contr_mm2_EZacc <- makeContrasts(da_treatmentR - da_treatmentC, da_treatmentRC - da_treatmentC, da_treatmentR - da_treatmentRC, levels = colnames(coef(vfit_mm2_EZacc)))

vfit_mm2_EZacc <- contrasts.fit(vfit_mm2_EZacc, contr_mm2_EZacc) #Estimate contrast for each gene
efit_mm2_EZacc <- eBayes(vfit_mm2_EZacc, robust = TRUE) #Empirical Bayes smoothing of standard errors

summary(decideTests(efit_mm2_EZacc, adjust.method="BH",p.value = 0.05))

```

# Separate countries for TDB EZacc
## Sweden

```{r Sweden TDB EZacc}

TDB_EZacc_SE <- order_TDB_EZacc[,c(59:88)]
dim(TDB_EZacc_SE) #55109    30

EZacc_SEdgelist <- DGEList(TDB_EZacc_SE)
dim(EZacc_SEdgelist) #55109    30

EZacc_SEdgelist <- calcNormFactors(EZacc_SEdgelist,method =c("TMM")) #trimmed mean of M-values

table(rowSums(EZacc_SEdgelist$counts==0)==30)
#FALSE 38967, TRUE 16142

#automatic way to filter genes
keep.exprs_EZaccSE <- filterByExpr(EZacc_SEdgelist)

#number of genes left after filtering
EZacc_SEdgelist <- EZacc_SEdgelist[keep.exprs_EZaccSE,, keep.lib.sizes=FALSE]
dim(EZacc_SEdgelist) #1548   30
sum(EZacc_SEdgelist$counts) #5302124

plotMDS(EZacc_SEdgelist, labels=treatmentSE)

all(rownames(EZfamSE_names) == colnames(EZacc_SEdgelist))

design_SEtreat_EZacc <- model.matrix(~0+treatmentSE) #if you remove 0 then it takes treatment C as intercept and serve as control
colnames(design_SEtreat_EZacc)

#voom model: mean-variance trend
v_SEtreat_EZacc <- voom(EZacc_SEdgelist, design_SEtreat_EZacc, plot=TRUE)

#final model:mean-variance trend
vfit_SEtreat_EZacc <- lmFit(v_SEtreat_EZacc, design_SEtreat_EZacc)
head(coef(vfit_SEtreat_EZacc))

#contrast
contr_SEtreat_EZacc <- makeContrasts(treatmentSERC - treatmentSER, treatmentSEC - treatmentSER, treatmentSEC - treatmentSERC, levels = colnames(coef(vfit_SEtreat_EZacc)))

vfit_SEtreat_EZacc <- contrasts.fit(vfit_SEtreat_EZacc, contr_SEtreat_EZacc)
efit_SEtreat_EZacc <- eBayes(vfit_SEtreat_EZacc, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_SEtreat_EZacc, main="Final model: Mean-variance trend")

summary(decideTests(efit_SEtreat_EZacc, adjust.method="BH",p.value = 0.05)) #all non-significant

##R vs RC+C

design_SE_drought_EZacc <- model.matrix(~drought_control_SE)
colnames(design_SE_drought_EZacc)

#voom model: mean-variance trend
v_SEdrought_EZacc <- voom(EZacc_SEdgelist, design_SE_drought_EZacc, plot=TRUE)

#final model:mean-variance trend
vfit_SEdrought_EZacc <- lmFit(v_SEdrought_EZacc, design_SE_drought_EZacc)
head(coef(vfit_SEdrought_EZacc))

efit_SEdrought_EZacc <- eBayes(vfit_SEdrought_EZacc, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_SEdrought_EZacc, main="Final model: Mean-variance trend")

summary(decideTests(efit_SEdrought_EZacc, adjust.method="BH", p.value=0.05))

```

## Germany

```{r Germany TDB EZacc}

TDB_EZacc_DE <- order_TDB_EZacc[,c(1:30)]
dim(TDB_EZacc_DE) #55109    30

EZacc_DEdgelist <- DGEList(TDB_EZacc_DE)
dim(EZacc_DEdgelist) #55109    30

EZacc_DEdgelist <- calcNormFactors(EZacc_DEdgelist,method =c("TMM")) #trimmed mean of M-values

table(rowSums(EZacc_DEdgelist$counts==0)==30)
#FALSE 38528, TRUE 16581 

#automatic way to filter genes
keep.exprs_EZaccDE <- filterByExpr(EZacc_DEdgelist)

#number of genes left after filtering
EZacc_DEdgelist <- EZacc_DEdgelist[keep.exprs_EZaccDE,, keep.lib.sizes=FALSE]
dim(EZacc_DEdgelist) #1440    30
sum(EZacc_DEdgelist$counts) #4629070

plotMDS(EZacc_DEdgelist, labels=treatmentDE)

all(rownames(EZfamDE_names) == colnames(EZacc_DEdgelist))

design_DEtreat_EZacc <- model.matrix(~0+treatmentDE) #if you remove 0 then it takes treatment C as intercept and serve as control
colnames(design_DEtreat_EZacc)

#voom model: mean-variance trend
v_DEtreat_EZacc <- voom(EZacc_DEdgelist, design_DEtreat_EZacc, plot=TRUE)

#final model:mean-variance trend
vfit_DEtreat_EZacc <- lmFit(v_DEtreat_EZacc, design_DEtreat_EZacc)
head(coef(vfit_DEtreat_EZacc))

#contrast
contr_DEtreat_EZacc <- makeContrasts(treatmentDERC - treatmentDER, treatmentDEC - treatmentDER, treatmentDEC - treatmentDERC, levels = colnames(coef(vfit_DEtreat_EZacc)))

vfit_DEtreat_EZacc <- contrasts.fit(vfit_DEtreat_EZacc, contr_DEtreat_EZacc)
efit_DEtreat_EZacc <- eBayes(vfit_DEtreat_EZacc, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_DEtreat_EZacc, main="Final model: Mean-variance trend")

summary(decideTests(efit_DEtreat_EZacc, adjust.method="BH",p.value = 0.05)) #1 down regulated
topTable(efit_DEtreat_EZacc, coef=2, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")
#ADJ45077.1

##R vs RC+C

design_DE_drought_EZacc <- model.matrix(~drought_control_DE)
colnames(design_DE_drought_EZacc)

#voom model: mean-variance trend
v_DEdrought_EZacc <- voom(EZacc_DEdgelist, design_DE_drought_EZacc, plot=TRUE)

#final model:mean-variance trend
vfit_DEdrought_EZacc <- lmFit(v_DEdrought_EZacc, design_DE_drought_EZacc)
head(coef(vfit_DEdrought_EZacc))

efit_DEdrought_EZacc <- eBayes(vfit_DEdrought_EZacc, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_DEdrought_EZacc, main="Final model: Mean-variance trend")

summary(decideTests(efit_DEdrought_EZacc, adjust.method="BH", p.value=0.05))
topTable(efit_DEdrought_EZacc, coef=2, number=10, adjust.method="BH", p.value=0.05, sort.by = "Amean", resort.by = "M")
#ADJ45077.1, ALX04048.1

```

## Spain

```{r Spain TDB EZacc}

TDB_EZacc_ES <- order_TDB_EZacc[,c(31:58)]
dim(TDB_EZacc_ES) #55109    28

EZacc_ESdgelist <- DGEList(TDB_EZacc_ES)
dim(EZacc_ESdgelist) #55109    28

EZacc_ESdgelist <- calcNormFactors(EZacc_ESdgelist,method =c("TMM")) #trimmed mean of M-values

table(rowSums(EZacc_ESdgelist$counts==0)==28)
#FALSE 37938, TRUE 17171 

#automatic way to filter genes
keep.exprs_EZaccES <- filterByExpr(EZacc_ESdgelist)

#number of genes left after filtering
EZacc_ESdgelist <- EZacc_ESdgelist[keep.exprs_EZaccES,, keep.lib.sizes=FALSE]
dim(EZacc_ESdgelist) #2443   28
sum(EZacc_ESdgelist$counts) #10821364

plotMDS(EZacc_ESdgelist, labels=treatmentES)

all(rownames(EZfamES_names) == colnames(EZacc_ESdgelist))

design_EStreat_EZacc <- model.matrix(~0+treatmentES) #if you remove 0 then it takes treatment C as intercept and serve as control
colnames(design_EStreat_EZacc)

#voom model: mean-variance trend
v_EStreat_EZacc <- voom(EZacc_ESdgelist, design_EStreat_EZacc, plot=TRUE)

#final model:mean-variance trend
vfit_EStreat_EZacc <- lmFit(v_EStreat_EZacc, design_EStreat_EZacc)
head(coef(vfit_EStreat_EZacc))

#contrast
contr_EStreat_EZacc <- makeContrasts(treatmentESRC - treatmentESR, treatmentESC - treatmentESR, treatmentESC - treatmentESRC, levels = colnames(coef(vfit_EStreat_EZacc)))

vfit_EStreat_EZacc <- contrasts.fit(vfit_EStreat_EZacc, contr_EStreat_EZacc)
efit_EStreat_EZacc <- eBayes(vfit_EStreat_EZacc, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_EStreat_EZacc, main="Final model: Mean-variance trend")

summary(decideTests(efit_EStreat_EZacc, adjust.method="BH",p.value = 0.05)) 

##R vs RC+C

design_ES_drought_EZacc <- model.matrix(~drought_control_ES)
colnames(design_ES_drought_EZacc)

#voom model: mean-variance trend
v_ESdrought_EZacc <- voom(EZacc_ESdgelist, design_ES_drought_EZacc, plot=TRUE)

#final model:mean-variance trend
vfit_ESdrought_EZacc <- lmFit(v_ESdrought_EZacc, design_ES_drought_EZacc)
head(coef(vfit_ESdrought_EZacc))

efit_ESdrought_EZacc <- eBayes(vfit_ESdrought_EZacc, robust = TRUE) #Empirical Bayes smoothing of standard errors
plotSA(efit_ESdrought_EZacc, main="Final model: Mean-variance trend")

summary(decideTests(efit_ESdrought_EZacc, adjust.method="BH", p.value=0.05))

```

```{r session}
session_info()
```