---
title: "TDB_relaxed functional composition"
author: "Katja Kozjek"
output: 
  rmarkdown::html_document:
    code_folding: 'hide'
    toc: true
    toc_float: true
    smart: true
    number_sections: false
    highlight: tango
    self_contained: true
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r wrap-hook, include=FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

# Load libraries

```{r load libraries, message=F, warning=F}
library(rafalib) # nice plot arrangement
library(phyloseq)
library(ranacapa) #for rarefaction curves
library(ggplot2)
library(wesanderson)
library(DESeq2)
library(vegan)
library(ggordiplots)
library(scales)
library(dplyr)
library(tibble)
library(devtools)
library(pairwiseAdonis)
library(limma)
library(cowplot)
library(tidyverse)
```

Analyses related to functional composition. Counts are filtered using a function filterByExpr. PCA and PERMANOVA are conducted with and without batch correction approach. For overall composition, batch correction is on countries, for each country separately batch correction is on field pair. Envfit added to the overall functional composition, soil and plant parameters are pre-selected. 

# Upload files

```{r load files, message=FALSE, warning=FALSE}

tdb_fam <- read.csv("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/TDB_EZfam_rem_filt.counts.tsv", sep = "\t", header = T, row.names = 1)

tdb_acc <- read.csv("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/TDB_EZacc_rem_filt.counts.tsv",sep = "\t", header = T, row.names = 1)

metafile = read.csv("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/metafile_relaxed.csv", sep=",", row.names=1)

metafile_remove <- metafile[-c(58,60),]

metafile_pca <- read.csv("C:/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/metafile_relaxed.csv", sep=",")
metafile_pca <- metafile_pca[-c(58,60),] #for pca dataframe

sampledata = sample_data(metafile)
sampledata_remove <- sampledata[-c(58,60),]

```

# PCA overall TDB EZfam
+ using counts filterByExpr
+ 2 Spanish samples are excluded

```{r PCA EZfam overall, message=FALSE, warning=FALSE}

dds_fam <- DESeqDataSetFromMatrix(countData = tdb_fam,
                              colData = metafile_remove,
                              design= ~ country + treatment)

vsd_fam <- varianceStabilizingTransformation(dds_fam)
vsd_fam_counts <- assay(vsd_fam) 

p1_EZfam_country <- plotPCA(vsd_fam, "country")
p2_EZfam_treat <- plotPCA(vsd_fam, "treatment")
plot_grid(p1_EZfam_country, p2_EZfam_treat)

#PERMANOVA

vsd_fam_counts <- t(vsd_fam_counts)
set.seed(4364)
adonis2(vsd_fam_counts ~ country*treatment, 
       data=metafile_remove, strata=metafile_remove$plot, permutations=999)

```
***EZfam level, country effect***

## Nice PCA EZfam

```{r PCA EZfam, message=FALSE, warning=FALSE}

#variance of components 
frac_var <- function(x) x^2/sum(x^2)

vsd_fam_pca <- prcomp(vsd_fam_counts)

vsd_fam_pca$sdev %>% 
  as_tibble() %>% 
  frac_var() %>% 
  mutate(Comp = colnames(vsd_fam_pca$x)) %>% 
  ggplot(aes(x=Comp, y = value)) + 
  geom_bar(stat = "identity", fill = "#4DC5F9") +
  geom_hline(yintercept = 0.03, linetype=2) +
  xlab("Principal Components") +
  scale_y_continuous(name = "Variance Explained", breaks = seq(0,0.8,0.1), 
                     labels =   percent_format(accuracy = 5L)) +
  theme_classic(base_size = 14)

vsd_fam_dframe <- vsd_fam_pca$x %>%
  as.data.frame() %>%
  rownames_to_column(var = "sample_ID") %>%
  full_join(metafile_pca, by = "sample_ID")

vsd_fam_dframe$treatment <- factor(vsd_fam_dframe$treatment,
                                        levels = c("C", "R", "RC"),
                                        labels = c("Control", "Roof", "Roof-Control"))

vsd_fam_dframe$country <- factor(vsd_fam_dframe$country,
                                      levels = c("SE", "DE", "ES"),
                                      labels = c("Sweden", "Germany", "Spain"))
#preselection for envfit 
sampledata_env <- sampledata_remove[,-c(8,11,13,14,16,17,18,19,20,23,25)]

envfit_pca_EZfam_all <- gg_envfit(vsd_fam_pca, sampledata_env, perm = 9999, scaling=1, plot=F)
envfit_pca_EZfam_all$df_arrows

pcaEZfam_overall <- ggplot(vsd_fam_dframe, aes(x=PC1, y=PC2, color = country)) +
  geom_point(size = 4,alpha=0.8) +
  geom_vline(xintercept = 0, linetype=2) +
  geom_hline(yintercept = 0, linetype=2) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) + labs(colour = "Country") + 
  scale_color_manual(values = c("#46ACC8", "#9C964A", "#B6854D"))

pcaEZfam_overall_env <- pcaEZfam_overall + geom_segment(data = envfit_pca_EZfam_all$df_arrows,
               aes(x = 0, xend = x, y = 0, yend = y), color = "black",
               arrow = arrow(length = unit(0.2,"cm"))) +
               geom_text(data = envfit_pca_EZfam_all$df_arrows,  
               aes(x = x, y = y, label = var), color = "black",
               size = 4, 
               hjust = -0.1) + xlab("PC1 (46.64%)") + ylab("PC2 (10.52%)")

print(pcaEZfam_overall_env)

#ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/pca_env_TDB_EZfam.jpeg", width = 30, height = 15, units = "cm")

```

### Batch correction on countries EZfam

```{r batch correction country EZfam, message=FALSE, warning=FALSE}

vsd_fam_remCOUNTRY <- vsd_fam
assay(vsd_fam_remCOUNTRY) <- limma::removeBatchEffect(assay(vsd_fam_remCOUNTRY), vsd_fam_remCOUNTRY$country)

p1_EZfam_country_corr <- plotPCA(vsd_fam_remCOUNTRY, "country")
p2_EZfam_treat_corr <- plotPCA(vsd_fam_remCOUNTRY, "treatment")
plot_grid(p1_EZfam_country_corr, p2_EZfam_treat_corr)

vsd_fam_remCOUNTRY_df <- as.data.frame(assay(vsd_fam_remCOUNTRY))
vsd_fam_remCOUNTRY_df <- t(vsd_fam_remCOUNTRY_df)

set.seed(4364)
adonis2(vsd_fam_remCOUNTRY_df ~ treatment, 
                     data=metafile_remove, strata=metafile_remove$plot, permutations=999) 

``` 
***No drought effect on overall functional composition EZfam after batch correction***

# PCA overall TDB EZacc
+ using counts filterByExpr
+ 2 Spanish samples are excluded

```{r PCA EZacc overall, message=FALSE, warning=FALSE}

dds_acc <- DESeqDataSetFromMatrix(countData = tdb_acc,
                              colData = metafile_remove,
                              design= ~ country + treatment)

vsd_acc <- varianceStabilizingTransformation(dds_acc)
vsd_acc_counts <- assay(vsd_acc)

p1_EZacc_country <- plotPCA(vsd_acc, "country")
p2_EZfam_treat <- plotPCA(vsd_acc, "treatment")
plot_grid(p1_EZacc_country, p2_EZfam_treat)

#PERMANOVA

vsd_acc_counts <- t(vsd_acc_counts)
set.seed(4364)
adonis2(vsd_acc_counts ~ country*treatment, 
       data=metafile_remove, strata=metafile_remove$plot, permutations=999)

```
***EZacc level, country effect***

## Nice PCA EZacc

```{r PCA EZacc, message=FALSE,warning=FALSE}

vsd_acc_pca <- prcomp(vsd_acc_counts)

vsd_acc_pca$sdev %>% 
  as_tibble() %>% 
  frac_var() %>% 
  mutate(Comp = colnames(vsd_acc_pca$x)) %>% 
  ggplot(aes(x=Comp, y = value)) + 
  geom_bar(stat = "identity", fill = "#4DC5F9") +
  geom_hline(yintercept = 0.03, linetype=2) +
  xlab("Principal Components") +
  scale_y_continuous(name = "Variance Explained", breaks = seq(0,0.8,0.1), 
                     labels =   percent_format(accuracy = 5L)) +
  theme_classic(base_size = 14)

vsd_acc_dframe <- vsd_acc_pca$x %>%
  as.data.frame() %>%
  rownames_to_column(var = "sample_ID") %>%
  full_join(metafile_pca, by = "sample_ID")

vsd_acc_dframe$treatment <- factor(vsd_acc_dframe$treatment,
                                        levels = c("C", "R", "RC"),
                                        labels = c("Control", "Roof", "Roof-Control"))

vsd_acc_dframe$country <- factor(vsd_acc_dframe$country,
                                      levels = c("SE", "DE", "ES"),
                                      labels = c("Sweden", "Germany", "Spain"))

envfit_pca_EZacc_all <- gg_envfit(vsd_acc_pca, sampledata_env, perm = 9999, scaling=1, plot=F)
envfit_pca_EZacc_all$df_arrows

pcaEZacc_overall <- ggplot(vsd_acc_dframe, aes(x=PC1, y=PC2, color = country)) +
  geom_point(size = 4,alpha=0.8) +
  geom_vline(xintercept = 0, linetype=2) +
  geom_hline(yintercept = 0, linetype=2) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) + labs(colour = "Country") + 
  scale_color_manual(values = c("#46ACC8", "#9C964A", "#B6854D"))

pcaEZacc_overall_env <- pcaEZacc_overall + geom_segment(data = envfit_pca_EZacc_all$df_arrows,
               aes(x = 0, xend = x, y = 0, yend = y), color = "black",
               arrow = arrow(length = unit(0.2,"cm"))) +
               geom_text(data = envfit_pca_EZacc_all$df_arrows,  
               aes(x = x, y = y, label = var), color = "black",
               size = 4, 
               hjust = -0.1) + xlab("PC1 (30.79%)") + ylab("PC2 (12.47%)")

print(pcaEZacc_overall_env)

#ggsave("/Users/Katja/Box Sync/PhD Lund Katja/Sequencing/Functional_genes/Diamond/data_TDB/relaxed/results/pca_env_TDB_EZacc.jpeg", width = 30, height = 15, units = "cm")

```

### Batch correction on countries EZacc

```{r batch correction country EZacc, message=FALSE,warning=FALSE}

vsd_acc_remCOUNTRY <- vsd_acc
assay(vsd_acc_remCOUNTRY) <- limma::removeBatchEffect(assay(vsd_acc_remCOUNTRY), vsd_acc_remCOUNTRY$country)

p1_EZacc_country_corr <- plotPCA(vsd_acc_remCOUNTRY, "country")
p2_EZacc_treat_corr <- plotPCA(vsd_acc_remCOUNTRY, "treatment")
plot_grid(p1_EZacc_country_corr, p2_EZacc_treat_corr)

vsd_acc_remCOUNTRY_df <- as.data.frame(assay(vsd_acc_remCOUNTRY))
vsd_acc_remCOUNTRY_df <- t(vsd_acc_remCOUNTRY_df)

#PERMANOVA 

set.seed(4364)
adonis2(vsd_acc_remCOUNTRY_df ~ treatment, 
                     data=metafile_remove, strata=metafile_remove$plot, permutations=999) 

``` 
***No drought effect on overall functional composition EZacc after batch correction***

# Separate countries
# Sweden 

## EZfam level

```{r SE family level, message=F, warning=F}

se_metadat <- metafile_remove %>%
  filter(country == "SE")

se_tdb_fam <- tdb_fam[,colnames(tdb_fam) %in% rownames(se_metadat)]

se_dds_fam <- DESeqDataSetFromMatrix(countData = se_tdb_fam,
                              colData = se_metadat[match(colnames(se_tdb_fam), rownames(se_metadat)), ],
                              design= ~ field_pair + treatment)

se_vsd_fam <- varianceStabilizingTransformation(se_dds_fam)
se_vsd_fam_counts <- assay(se_vsd_fam)

p1_SE_FP <- plotPCA(se_vsd_fam, "field_pair")
p2_SE_treat <- plotPCA(se_vsd_fam, "treatment")
plot_grid(p1_SE_FP, p2_SE_treat)

#PERMANOVA
se_vsd_fam_counts <- t(se_vsd_fam_counts)
set.seed(4364)
adonis2(se_vsd_fam_counts~ treatment, 
       data=se_metadat, strata=se_metadat$field_pair, permutations=999)

```
***No drought effect on family level***

### batch correction 

```{r SE batch correction, family level, message=F, warning=F}

se_vsd_fam_remFP <- se_vsd_fam
assay(se_vsd_fam_remFP) <- limma::removeBatchEffect(assay(se_vsd_fam_remFP), se_vsd_fam_remFP$field_pair)

p1_SE_FP_corr <- plotPCA(se_vsd_fam_remFP, "field_pair")
p2_SE_treat_corr <- plotPCA(se_vsd_fam_remFP, "treatment")
plot_grid(p1_SE_FP_corr, p2_SE_treat_corr)

#PERMANOVA 
se_vsd_fam_remFP_df <- as.data.frame(assay(se_vsd_fam_remFP))
se_vsd_fam_remFP_df <- t(se_vsd_fam_remFP_df)

set.seed(4364)
adonis2(se_vsd_fam_remFP_df ~ treatment, 
                     data=se_metadat, permutations=999)

```
***No drought effect on family level after batch correction on field pairs***

## EZ ID level

```{r SE ID level, message=F, warning=F}

se_tdb_acc <- tdb_acc[,colnames(tdb_acc) %in% rownames(se_metadat)]

se_dds_acc <- DESeqDataSetFromMatrix(countData = se_tdb_acc,
                              colData = se_metadat[match(colnames(se_tdb_acc), rownames(se_metadat)), ],
                              design= ~ field_pair + treatment)

se_vsd_acc <- varianceStabilizingTransformation(se_dds_acc)
se_vsd_acc_counts <- assay(se_vsd_acc)

p1_SEacc_FP <- plotPCA(se_vsd_acc, "field_pair")
p2_SEacc_treat <- plotPCA(se_vsd_acc, "treatment")
plot_grid(p1_SEacc_FP, p2_SEacc_treat)

#PERMANOVA
se_vsd_acc_counts <- t(se_vsd_acc_counts)
set.seed(4364)
adonis2(se_vsd_acc_counts ~ treatment, 
       data=se_metadat, strata=se_metadat$field_pair, permutations=999)

```
***No drought effect on ID level***

### batch correction 

```{r SE batch correction, ID level, message=F, warning=F}

se_vsd_acc_remFP <- se_vsd_acc
assay(se_vsd_acc_remFP) <- limma::removeBatchEffect(assay(se_vsd_acc_remFP), se_vsd_acc_remFP$field_pair)

p1_SEacc_FP_corr <- plotPCA(se_vsd_acc_remFP, "field_pair")
p2_SEacc_treat_corr <- plotPCA(se_vsd_acc_remFP, "treatment")
plot_grid(p1_SEacc_FP_corr, p2_SEacc_treat_corr)

#PERMANOVA 
se_vsd_acc_remFP_df <- as.data.frame(assay(se_vsd_acc_remFP))
se_vsd_acc_remFP_df <- t(se_vsd_acc_remFP_df)

set.seed(4364)
adonis2(se_vsd_acc_remFP_df ~ treatment, 
                     data=se_metadat, permutations=999)

```
***No drought effect on ID level after batch correction***

# Germany

## EZfam level

```{r DE family level, message=F, warning=F}

de_metadat <- metafile_remove %>%
  filter(country == "DE")

de_tdb_fam <- tdb_fam[,colnames(tdb_fam) %in% rownames(de_metadat)]

de_dds_fam <- DESeqDataSetFromMatrix(countData = de_tdb_fam,
                              colData = de_metadat[match(colnames(de_tdb_fam), rownames(de_metadat)), ],
                              design= ~ field_pair + treatment)

de_vsd_fam <- varianceStabilizingTransformation(de_dds_fam)
de_vsd_fam_counts <- assay(de_vsd_fam)

p1_DE_FP <- plotPCA(de_vsd_fam, "field_pair")
p2_DE_treat <- plotPCA(de_vsd_fam, "treatment")
plot_grid(p1_DE_FP, p2_DE_treat)

#PERMANOVA
de_vsd_fam_counts <- t(de_vsd_fam_counts)
set.seed(4364)
adonis2(de_vsd_fam_counts~ treatment, 
       data=de_metadat, strata=de_metadat$field_pair, permutations=999)

```
***No drought effect on family level***

### batch correction 

```{r DE batch correction, family level, message=F, warning=F}

de_vsd_fam_remFP <- de_vsd_fam
assay(de_vsd_fam_remFP) <- limma::removeBatchEffect(assay(de_vsd_fam_remFP), de_vsd_fam_remFP$field_pair)

p1_DE_FP_corr <- plotPCA(de_vsd_fam_remFP, "field_pair")
p2_DE_treat_corr <- plotPCA(de_vsd_fam_remFP, "treatment")
plot_grid(p1_DE_FP_corr, p2_DE_treat_corr)

#PERMANOVA 
de_vsd_fam_remFP_df <- as.data.frame(assay(de_vsd_fam_remFP))
de_vsd_fam_remFP_df <- t(de_vsd_fam_remFP_df)

set.seed(4364)
adonis2(de_vsd_fam_remFP_df ~ treatment, 
                     data=de_metadat, permutations=999)

```
***Slight drought effect on family level after batch correction***

## EZ ID level

```{r DE ID level, message=F, warning=F}

de_tdb_acc <- tdb_acc[,colnames(tdb_acc) %in% rownames(de_metadat)]

de_dds_acc <- DESeqDataSetFromMatrix(countData = de_tdb_acc,
                              colData = de_metadat[match(colnames(de_tdb_acc), rownames(de_metadat)), ],
                              design= ~ field_pair + treatment)

de_vsd_acc <- varianceStabilizingTransformation(de_dds_acc)
de_vsd_acc_counts <- assay(de_vsd_acc)

p1_DEacc_FP <- plotPCA(de_vsd_acc, "field_pair")
p2_DEacc_treat <- plotPCA(de_vsd_acc, "treatment")
plot_grid(p1_DEacc_FP, p2_DEacc_treat)

#PERMANOVA 
de_vsd_acc_counts <- t(de_vsd_acc_counts)
set.seed(4364)
adonis2(de_vsd_acc_counts ~ treatment, 
                     data=de_metadat, strata=de_metadat$field_pair, permutations=999)

```
***No drought effect on ID level***

### batch correction 

```{r DE batch correction, ID level, message=F, warning=F}

de_vsd_acc_remFP <- de_vsd_acc
assay(de_vsd_acc_remFP) <- limma::removeBatchEffect(assay(de_vsd_acc_remFP), de_vsd_acc_remFP$field_pair)

p1_DEacc_FP_corr <- plotPCA(de_vsd_acc_remFP, "field_pair")
p2_DEacc_treat_corr <- plotPCA(de_vsd_acc_remFP, "treatment")
plot_grid(p1_DEacc_FP_corr, p2_DEacc_treat_corr)

#PERMANOVA 
de_vsd_acc_remFP_df <- as.data.frame(assay(de_vsd_acc_remFP))
de_vsd_acc_remFP_df <- t(de_vsd_acc_remFP_df)

set.seed(4364)
adonis2(de_vsd_acc_remFP_df ~ treatment, 
                     data=de_metadat, permutations=999)

```
***Slight drought effect on ID level after batch correction***

# Spain

## EZfam level

```{r ES family level, message=F, warning=F}

es_metadat <- metafile_remove %>%
  filter(country == "ES")

es_tdb_fam <- tdb_fam[,colnames(tdb_fam) %in% rownames(es_metadat)]

es_dds_fam <- DESeqDataSetFromMatrix(countData = es_tdb_fam,
                              colData = es_metadat[match(colnames(es_tdb_fam), rownames(es_metadat)), ],
                              design= ~ field_pair + treatment)

es_vsd_fam <- varianceStabilizingTransformation(es_dds_fam)
es_vsd_fam_counts <- assay(es_vsd_fam)

p1_ES_EP <- plotPCA(es_vsd_fam, "field_pair")
p2_ES_treat <- plotPCA(es_vsd_fam, "treatment")
plot_grid(p1_ES_EP, p2_ES_treat)

#PERMANOVA
es_vsd_fam_counts <- t(es_vsd_fam_counts)
set.seed(4364)
adonis2(es_vsd_fam_counts ~ treatment, 
       data=es_metadat, strata=es_metadat$field_pair, permutations=999)

```
***No drought on family level***

### batch correction 

```{r ES batch correction, family level, message=F, warning=F}

es_vsd_fam_remFP <- es_vsd_fam
assay(es_vsd_fam_remFP) <- limma::removeBatchEffect(assay(es_vsd_fam_remFP), es_vsd_fam_remFP$field_pair)

p1_ES_EP_corr <- plotPCA(es_vsd_fam_remFP, "field_pair")
p2_ES_treat_corr <- plotPCA(es_vsd_fam_remFP, "treatment")
plot_grid(p1_ES_EP_corr, p2_ES_treat_corr)

#PERMANOVA 
es_vsd_fam_remFP_df <- as.data.frame(assay(es_vsd_fam_remFP))
es_vsd_fam_remFP_df <- t(es_vsd_fam_remFP_df)

set.seed(4364)
adonis2(es_vsd_fam_remFP_df ~ treatment, 
                     data=es_metadat, permutations=999)

```
***No drought on family level after batch correction***

## EZ ID level

```{r ES ID level, message=F, warning=F}

es_tdb_acc <- tdb_acc[,colnames(tdb_acc) %in% rownames(es_metadat)]

es_dds_acc <- DESeqDataSetFromMatrix(countData = es_tdb_acc,
                              colData = es_metadat[match(colnames(es_tdb_acc), rownames(es_metadat)), ],
                              design= ~ field_pair + treatment)

es_vsd_acc <- varianceStabilizingTransformation(es_dds_acc)
es_vsd_acc_counts <- assay(es_vsd_acc)

p1_ESacc_EP <- plotPCA(es_vsd_acc, "field_pair")
p2_ESacc_treat <- plotPCA(es_vsd_acc, "treatment")
plot_grid(p1_ESacc_EP, p2_ESacc_treat)

#PERMANOVA 
es_vsd_acc_counts <- t(es_vsd_acc_counts)
set.seed(4364)
adonis2(es_vsd_acc_counts ~ treatment, 
                     data=es_metadat, strata=es_metadat$field_pair, permutations=999)

```
***No drought effect on ID level***

### batch correction 

```{r ES batch correction, ID level, message=F, warning=F}

es_vsd_acc_remFP <- es_vsd_acc
assay(es_vsd_acc_remFP) <- limma::removeBatchEffect(assay(es_vsd_acc_remFP), es_vsd_acc_remFP$field_pair)

p1_ESacc_EP_corr <- plotPCA(es_vsd_acc_remFP, "field_pair")
p2_ESacc_treat_corr <- plotPCA(es_vsd_acc_remFP, "treatment")
plot_grid(p1_ESacc_EP_corr, p2_ESacc_treat_corr)

#PERMANOVA 
es_vsd_acc_remFP_df <- as.data.frame(assay(es_vsd_acc_remFP))
es_vsd_acc_remFP_df <- t(es_vsd_acc_remFP_df)

set.seed(4364)
adonis2(es_vsd_acc_remFP_df ~ treatment, 
                     data=es_metadat, permutations=999)

```
***No drought effect on ID level after batch correction***

```{r session}

sessionInfo()

```
